---
title: "SOS"
author: "ST, YL, EH, AH"
date: "2023-11-26"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step 1. Install / load packages

```{r message = FALSE, warning=FALSE}
# Install and load required packages
library(purrr)
library(dplyr)
library(ggplot2)
library(ggfortify)
```

## Step 2. Read each .txt file into R as a data frame 

Ensure .Rmd file and all timepoint.txt files are in the same working directory

```{r warning=FALSE}

# List all timepoint.txt files in the directory
txt_files <- list.files(pattern = ".txt", full.names = FALSE)

# Empty list of data frame names
list_df_names <- list()

# Function to read a .txt file and return as a data frame
read_and_number_df <- function(file_path) {
  # Read the file into a data frame
  df <- read.table(file_path, sep = ",", header = FALSE, col.names = c("x_pixel", "y_pixel", "Frame_number", "Molecule_ID"))

  # Extract the file name
  file_name <- sub(pattern = ".txt", "", basename(file_path))
  
  # Assign file name to the data frame
  assign(paste0(file_name), df, envir = .GlobalEnv)
  
  # Append data frame name to list_df_names
  list_df_names <<- append(list_df_names, toString(file_name))
}

# Use purrr::walk to apply the function to each file in the txt_files list
walk(txt_files, read_and_number_df)

```

## Step 3. Calculate x and y physical lengths for each molecule 

```{r warning=FALSE}

# Function to calculate x and y physical length from pixels for each molecule 
calc_phys_length <- function(timepoints) {
  # Retrieve the data frame
  df <- get(timepoints)
  
  # Calculate physical lengths and create a new column "x_pos" and "y_pos"
  # Conversion rate between pixel and physical is 0.107μm (pixel)^(-1)
  df$x_pos <- df$x_pixel * 0.107
  df$y_pos <- df$y_pixel * 0.107
  
  # Assign the modified data frame back
  assign(timepoints, df, envir = .GlobalEnv)
}

# Use purr::walk to apply the function to each data frame in the list
walk(list_df_names, calc_phys_length)

```

## Step 4. Filter out cases where Molecule_ID frequency is <= 4

```{r warning=FALSE}

# Function to filter out the frequencies of Molecule_ID with 4 or less records
filter_data <- function(timepoints) {
  
  # Retrieve data frame
  df <- get(timepoints)
  
  # Convert the Molecule ID column to factor (categorical variable)
  df$Molecule_ID <- as.factor(df$Molecule_ID)
  
  # Filter to only include molecule IDs with 5 or more records
  df <- df %>% group_by(Molecule_ID) %>% 
    filter(n() >= 5) 
  
  # Return modified data frame back
  assign(timepoints, df, envir = .GlobalEnv)
}

# Use purr::walk to apply the function to each data frame in the list
walk(list_df_names, filter_data)

```

## Step 5. Check for consecutive frames for each Molecule_ID

```{r warning=FALSE} 

# Function to check for consecutive frames for each Molecule_ID 
check_frame <- function(timepoints) {
  result <- timepoints %>%
    group_by(Molecule_ID) %>%
    arrange(Frame_number) %>%
    filter(
      row_number() <= 5 &
      all(diff(Frame_number) == 1) &
      sum(diff(Frame_number) != 1) <= 1
    ) %>%
    ungroup()
  
  return(result)
}

# Return the results to the data frame
frame_to_df <- function(df_name) {
  # Retrieve data frame
  df <- get(df_name)
  
  # Apply the function to our data
  df_filtered <- check_frame(df)
  
  # Assign the filtered data frame back to the environment
  assign(df_name, df_filtered, envir = .GlobalEnv)
}
 
# Use purr::walk to apply the function to each data frame in the list
walk(list_df_names, frame_to_df)

```

## Step 6. Calculate mean squared difference (MSD) and diffusion coefficient (D)

```{r warning=FALSE}

# Function to Calculate mean squared difference (MSD) and diffusion coefficient (D)
calc_MSD_D <- function(timepoints){
  # Retrieve the data frame
  df <- get(timepoints)
  
  # Create new data frame
  df_new <- data.frame(matrix(ncol = 3, nrow = 0))
  
  # Assign name to new data frame based on old data frame
  df_new_name <- paste0(timepoints, "_calculations")

  # Convert the Molecule ID column to factor (categorical variable)
  df$Molecule_ID <- as.factor(df$Molecule_ID)
   
  N = 5     # Number of frames
  dT = 0.00748 # delta time (s)
  
  # Filter by molecule ID and calculate MSD and D 
  df_new <- df %>% group_by(Molecule_ID) %>% 
    summarize(MSD = (1/4)*sum((lead(x_pos, default = last(x_pos)) - x_pos)^2 +
    (lead(y_pos, default = last(y_pos)) - y_pos)^2)) %>% 
    mutate(D = MSD/(4*dT))
 
  # Return new data frame back to global environment
  assign(df_new_name, df_new, envir = .GlobalEnv)
  
  #Print progress 
  cat("Finished for", timepoints, "\n")
}    

# Use purr::walk to apply the function to each data frame in the list
walk(list_df_names, calc_MSD_D)

```
## Step 7. Generate a dictionary for each timepoint and antibiotic treatment time. 

Create a new list of data frames with calculated D.

```{r warning=FALSE}

# Dictionary for each timepoint and corresponding minutes after antibiotic treatment
timepoint_dictionary <- c("01"="0", "02"="5", "03"="10", "04"="16", "05"="22", 
  "06"="29", "07"="34", "08"="40", "09"="45", "10"="49", "11"="54", "12"="61", 
  "13"="67", "14"="74", "15"="80", "16"="87", "17"="94", "18"="101")

# List of data frames with calculated diffusion coefficients
list_timepoint_calc <- c('timepoint_01_calculations', 'timepoint_02_calculations', 'timepoint_03_calculations', 'timepoint_04_calculations', 'timepoint_05_calculations', 'timepoint_06_calculations', 'timepoint_07_calculations', 'timepoint_08_calculations', 'timepoint_09_calculations', 'timepoint_10_calculations', 'timepoint_11_calculations', 'timepoint_12_calculations', 'timepoint_13_calculations', 'timepoint_14_calculations', 'timepoint_15_calculations', 'timepoint_16_calculations', 'timepoint_17_calculations', 'timepoint_18_calculations')

```

## Step 8. Determine the state of each LexA molecule based on D 

```{r warning=FALSE}

# Function to determine LexA state based on D
LexA_state <- function(timepoints){
  # Retrieve the data frame
  df <- get(timepoints)

  # Determine LexA state
  df <- df %>% 
    mutate(state = case_when(D < 0.2 ~ 'Bound',
                           D < 2.0 ~ 'Free',
                           D > 2.0 ~ 'Degraded'))
  
  # Return modified data frame back
  assign(timepoints, df, envir = .GlobalEnv)
}

# Use purr::walk to apply the function to each data frame in the list
walk(list_timepoint_calc, LexA_state)

```

## Step 9. Generate histogram plots for each timepoint

```{r warning=FALSE}

# Empty list for plots
list_hist <- list()

# Iterate through list_timepoint_calc and generate plots
for (x in 1:length(list_timepoint_calc)){
  time <- timepoint_dictionary[as.numeric(gsub("\\D", "", list_timepoint_calc[x]))]
  
  # Generate histogram title
  title <- paste("Distribution of LexA Diffusion, t =", time, "min \n")
  
  #Get data frame
  df <- (get(list_timepoint_calc[x]))
  
  # Plot histogram
  hist <- ggplot(df, aes(x=D)) + geom_histogram(binwidth = .4, fill = "darkgrey", aes(y=..count../sum(..count..))) + 
    ylim(0,.1) + xlim(0,15) +
    labs(x = bquote('Diffusion Coefficient (D)' ~ (μm^2/s)), y = "Proportion", title = title) + 
    theme_classic(base_size = 15) + theme(plot.title = element_text(hjust = 0.5))
  
  # Store histogram in list_hist
  list_hist[[length(list_hist) + 1]] <- hist
  
  # Save all histograms as .png to working directory
  file_name <- paste("Distribution", time, "min.png")
  ggsave(file_name, height = 7, width = 8)
}

# Show all plots
for (plot in list_hist) {
  print(plot)
}

```

## Step 10. Obtain the descriptive stats of all the timepoints

```{r warning=FALSE}

library(pastecs)

list_df_calculations <- list(timepoint_01_calculations,timepoint_02_calculations,timepoint_03_calculations, timepoint_04_calculations, timepoint_05_calculations, timepoint_06_calculations, timepoint_07_calculations, timepoint_08_calculations, timepoint_09_calculations, timepoint_10_calculations, timepoint_11_calculations, timepoint_12_calculations, timepoint_13_calculations, timepoint_14_calculations, timepoint_15_calculations, timepoint_16_calculations, timepoint_17_calculations, timepoint_18_calculations)

# Function to calculate descriptive statistics for each data frame
calculate_stats <- function(df) {
  return(stat.desc(df[,2:3]))
}
 
# Use purrr::map to apply the function to each data frame in the list
stats_list <- purrr::map(list_df_calculations, calculate_stats)

# Print the results
print(stats_list)

View(stats_list)

```

## Step 11. Plot mean D against time

```{r warning=FALSE}
# Create data frame of mean and CI of D
D_stats <- setNames(data.frame(matrix(ncol = 4, nrow = 18)), c("Time (min)", "Mean D", "Upper D", "Lower D"))

# Loop through stats_list and populate D_stats 
for (i in 1:length(stats_list)){
  # Get time from dictionary
  time <- as.numeric(timepoint_dictionary[i])
  
  # Get mean value of D
  mean <- stats_list[[i]][[9,2]]

  # Get CI of D
  upper <- stats_list[[i]][[9,2]] + stats_list[[i]][[11,2]]
  lower <- stats_list[[i]][[9,2]] - stats_list[[i]][[11,2]]
  
  # Populate data frame
  D_stats[i,] <- c(time, mean, upper, lower)
  
  # Row 19 is control
}


# Plot graph of mean D against time, with control
ggplot(D_stats, aes(x = `Time (min)`, y = `Mean D`)) + geom_point() + 
  ylim(2,6) + 
  geom_line() + 
  geom_errorbar(aes(ymin = `Lower D`, ymax = `Upper D`), alpha = 0.3) + 
  labs(y = bquote('Mean Diffusion Coefficient' ~ (μm^2/s))) + 
  theme_classic(base_size = 15) + 
  geom_hline(yintercept = D_stats[19,2], linetype='dashed', col = 'darkred') +
  annotate("text", x = 100, y = 3.6, label = "Untreated", col = 'darkred')



# Save as .png to working directory
#ggsave("Mean_plot.png", height = 7 , width = 8)

```

## Step 12. Fit linear regression

Gather raw data across all timepoints

```{r warning=FALSE}

# Create new data frame
D_reg <- setNames(data.frame(matrix(ncol = 3)), c("Time (min)", "Molecule_ID", "D"))

# Function to generate data frame for linear regression
Lin_reg <- function(timepoints){
  # Retrieve the data frame
  df <- get(timepoints)
  
  # Get time from dictionary
  time <- timepoint_dictionary[as.numeric(gsub("\\D", "", timepoints))]
  
  # Add new time col to old data frame
  df$`Time (min)` <- time
  
  # Make sure the columns have the same types
  df$Molecule_ID <- as.character(df$Molecule_ID)
  df$D <- as.numeric(df$D)
  
  # Import cols to new data frame
  D_reg <- bind_rows(D_reg, df[, c("Time (min)", "Molecule_ID", "D")])
  
  # Return to global env
  assign("D_reg", D_reg, envir = .GlobalEnv)
} 

# Use purrr::walk to apply the function to each file in the data frame list
walk(list_timepoint_calc[1:18], Lin_reg)

# Delete first (empty) row
D_reg <- D_reg[-1, ]

```

Fit a linear model

```{r warning=FALSE}

# Ensure time is numeric class
D_reg[,1] <- as.numeric(D_reg$`Time (min)`)

# Fit lm
lm <- lm(D ~ `Time (min)`, data = D_reg)

# Display lm plots
plot(lm) 
autoplot(lm)

```

We can see here that the data is fairly linear, but also heteroscedastic and non-normal.
So, we fit a generalised linear model.

```{r warning=FALSE}

# Fit glm
glm <- glm(D ~ `Time (min)`, data = D_reg, family = Gamma(link = "inverse"))

# Display plots
autoplot(glm)
```

Superimpose the glm on the raw data and mean 

```{r warning=FALSE}

# Superimpose glm on raw points
ggplot(aes(x = `Time (min)`, y = D), data = D_reg) + geom_point(position = "jitter", alpha = .1, size = .75) + 
  stat_smooth(method = "glm", formula = y ~ x, method.args = list(family = Gamma(link = "log")), color = "blue") + 
  ylim(0,15) + 
  labs(y = bquote('Diffusion Coefficient' ~ (μm^2/s))) +
  theme_classic(base_size = 15)


# Superimpose glm on mean points
ggplot(D_stats, aes(x = `Time (min)`, y = `Mean D`)) + 
  geom_point() + 
  ylim(2,6) + geom_line() + 
  geom_errorbar(aes(ymin = `Lower D`, ymax = `Upper D`), alpha = 0.3) + 
  labs(title = '', y = bquote('Mean Diffusion Coefficient' ~ (μm^2/s))) + 
  theme_classic(base_size = 15) + 
  stat_smooth(method = "glm", formula = y ~ x, method.args = list(family = Gamma(link = "log")), color = "blue") +
  geom_hline(yintercept = D_stats[19,2], linetype='dashed', col = 'darkred') +
  annotate("text", x = 100, y = 3.6, label = "Untreated", col = 'darkred')


```

## Step 13. Compare the coefficient distribution of the treated (across time) and untreated population 

```{r warning=FALSE}
#Determine if the data have a normal distribution with the shapiro-wilk test.

#Test each timepoint calculation. If p is significant it means it is not normal

shapiro_01 <- shapiro.test(timepoint_01_calculations$D)
shapiro_02 <- shapiro.test(timepoint_02_calculations$D)
shapiro_03 <- shapiro.test(timepoint_03_calculations$D)
shapiro_04 <- shapiro.test(timepoint_04_calculations$D)
shapiro_05 <- shapiro.test(timepoint_05_calculations$D)
shapiro_06 <- shapiro.test(timepoint_06_calculations$D)
shapiro_07 <- shapiro.test(timepoint_07_calculations$D)
shapiro_08 <- shapiro.test(timepoint_08_calculations$D)
shapiro_09 <- shapiro.test(timepoint_09_calculations$D)
shapiro_10 <- shapiro.test(timepoint_10_calculations$D)
shapiro_11 <- shapiro.test(timepoint_11_calculations$D)
shapiro_12 <- shapiro.test(timepoint_12_calculations$D)
shapiro_13 <- shapiro.test(timepoint_13_calculations$D)
shapiro_14 <- shapiro.test(timepoint_14_calculations$D)
shapiro_15 <- shapiro.test(timepoint_15_calculations$D)
shapiro_16 <- shapiro.test(timepoint_16_calculations$D)
shapiro_17 <- shapiro.test(timepoint_17_calculations$D)
shapiro_18 <- shapiro.test(timepoint_18_calculations$D)


#Join all the tests in a single dataframe
normality_Dcoeff <- list(shapiro_01, shapiro_02,shapiro_03, shapiro_04, shapiro_05, shapiro_06, shapiro_07, shapiro_08, shapiro_09, shapiro_10, shapiro_11, shapiro_12, shapiro_13, shapiro_14, shapiro_15, shapiro_16, shapiro_17, shapiro_18)


#print(normality_Dcoeff)


#All the results show that the distributions are not normal

#Test for normality the untreated_tracks 

shapiro_control <- shapiro.test(untreated_tracks_calculations$D)
print(shapiro_control)
```


```{r warning=FALSE}
#Export the time point calculations as csv and join the D columns in a single file for the Mann Whitney test


write.csv(timepoint_01_calculations, 'timepoint_01_calculations.csv', row.names = FALSE)
write.csv(timepoint_02_calculations, 'timepoint_02_calculations.csv', row.names = FALSE)
write.csv(timepoint_03_calculations, 'timepoint_03_calculations.csv', row.names = FALSE)
write.csv(timepoint_04_calculations, 'timepoint_04_calculations.csv', row.names = FALSE)
write.csv(timepoint_05_calculations, 'timepoint_05_calculations.csv', row.names = FALSE)
write.csv(timepoint_06_calculations, 'timepoint_06_calculations.csv', row.names = FALSE)
write.csv(timepoint_07_calculations, 'timepoint_07_calculations.csv', row.names = FALSE)
write.csv(timepoint_08_calculations, 'timepoint_08_calculations.csv', row.names = FALSE)
write.csv(timepoint_09_calculations, 'timepoint_09_calculations.csv', row.names = FALSE)
write.csv(timepoint_10_calculations, 'timepoint_10_calculations.csv', row.names = FALSE)
write.csv(timepoint_11_calculations, 'timepoint_11_calculations.csv', row.names = FALSE)
write.csv(timepoint_12_calculations, 'timepoint_12_calculations.csv', row.names = FALSE)
write.csv(timepoint_13_calculations, 'timepoint_13_calculations.csv', row.names = FALSE)
write.csv(timepoint_14_calculations, 'timepoint_14_calculations.csv', row.names = FALSE)
write.csv(timepoint_15_calculations, 'timepoint_15_calculations.csv', row.names = FALSE)
write.csv(timepoint_16_calculations, 'timepoint_16_calculations.csv', row.names = FALSE)
write.csv(timepoint_17_calculations, 'timepoint_17_calculations.csv', row.names = FALSE)
write.csv(timepoint_18_calculations, 'timepoint_18_calculations.csv', row.names = FALSE)
write.csv(timepoint_10_calculations, 'timepoint_10_calculations.csv', row.names = FALSE)
write.csv(untreated_tracks_calculations, 'untreated_tracks_calculations.csv', row.names = FALSE)

```



```{r warning=FALSE}

#Import the csv file that contains only the diffussion coefficients of the timepoints and the untreated population

timepoints_coeff <- read.csv("D_coefficients.csv", header = TRUE)


colnames(timepoints_coeff)
```


```{r warning=FALSE}
#Conduct comparison for non-normal populations using the Mann Whitney test (wilcox.test)

mw_01 <- wilcox.test(timepoints_coeff$timepoint01_D, timepoints_coeff$untreated)
mw_01

mw_02 <- wilcox.test(timepoints_coeff$timepoint02_D, timepoints_coeff$untreated)
mw_02


mw_03 <- wilcox.test(timepoints_coeff$timepoint03_D, timepoints_coeff$untreated)
mw_03


mw_04 <- wilcox.test(timepoints_coeff$timepoint04_D, timepoints_coeff$untreated)
mw_04

mw_05 <- wilcox.test(timepoints_coeff$timepoint05_D, timepoints_coeff$untreated)
mw_05

mw_06 <- wilcox.test(timepoints_coeff$timepoint06_D, timepoints_coeff$untreated)
mw_06

mw_07 <- wilcox.test(timepoints_coeff$timepoint07_D, timepoints_coeff$untreated)
mw_07


mw_08 <- wilcox.test(timepoints_coeff$timepoint08_D, timepoints_coeff$untreated)
mw_08

mw_09 <- wilcox.test(timepoints_coeff$timepoint09_D, timepoints_coeff$untreated)
mw_09


mw_10 <- wilcox.test(timepoints_coeff$timepoint10_D, timepoints_coeff$untreated)
mw_10

mw_11 <- wilcox.test(timepoints_coeff$timepoint11_D, timepoints_coeff$untreated)
mw_11


mw_12 <- wilcox.test(timepoints_coeff$timepoint12_D, timepoints_coeff$untreated)
mw_12

mw_13 <- wilcox.test(timepoints_coeff$timepoint13_D, timepoints_coeff$untreated)
mw_13

mw_14 <- wilcox.test(timepoints_coeff$timepoint14_D, timepoints_coeff$untreated)
mw_14


mw_15 <- wilcox.test(timepoints_coeff$timepoint15_D, timepoints_coeff$untreated)
mw_15

mw_16 <- wilcox.test(timepoints_coeff$timepoint16_D, timepoints_coeff$untreated)
mw_16

mw_17 <- wilcox.test(timepoints_coeff$timepoint17_D, timepoints_coeff$untreated)
mw_17

mw_18 <- wilcox.test(timepoints_coeff$timepoint18_D, timepoints_coeff$untreated)
mw_18
```

